### 2024-11-03 teaching/math240f24/draft.tex
hallowe'en lecture

We now define certain important families of graphs. For any integer $n\ge 1$, we define
the {\it complete graph} $K_n$ to be the graph on $n$ vertices with every possible edge present.
Small examples are illustrated in Fig.~2.
\midinsert
$$\epsfbox{kn.ps}$$
\vskip5pt
\caption{Complete graphs $K_n$ for small $n$.}
\endinsert
\goodbreak
\bye
For an integer $n\ge 1$,
we define the {\it Hamming cube} or {\it hypercube} to be the graph $Q_n$
whose vertex set is $\{0,1\}^n$, the set of all binary strings of length $n$,
and whose edge set is the exactly the set of pairs of strings that differ by exactly
one bit. Small examples are illustrated in Fig.~3. (As an exercise, try to figure out
which strings correspond to which vertices!)
\midinsert
$$\epsfbox{qn.ps}$$
\vskip5pt
\caption{Hypercubes $Q_n$ for small $n$.}
\endinsert
\goodbreak

@@

\medskip\boldlabel\datestamp{18}{} Walks, paths, and cycles.
A {\it walk} in $G = (V,E)$ is a sequence $\sigma$ of vertices
$$\sigma = (v_0, v_1, \ldots, v_n)$$
such that $v_i v_{i+1}\in E$ for all $0\le i \le n-1$. The {\it endpoints}
of $\sigma$ are $v_0$ and $v_n$, and the {\it length} of $\sigma$, denoted $|\sigma|$,
is the number of edges traversed, namely, $n$.
The walk is said to be {\it closed} if $v_0 = v_n$ and {\it open} otherwise.
A walk is called a {\it path} if no vertex is repeated.

@@

We conclude that $\sigma$ is a cycle.\slug

@@

Another characterisation of trees is that they are exactly the connected graphs with $n-1$
nodes. To prove this, we first need a lemma.

>>>>

We now define certain important families of graphs. For any integer $n\ge 1$, we define
the {\it complete graph} $K_n$ to be the graph on $n$ vertices with every possible edge present.
Small examples are illustrated in Fig.~2.
\midinsert
$$\epsfbox{kn.ps}$$
\vskip5pt
\caption{Complete graphs $K_n$ for small $n$.}
\endinsert
\goodbreak
\noindent\datestamp{31}{x} \hskip\parindent For an integer $n\ge 1$,
we define the {\it Hamming cube} or {\it hypercube} to be the graph $Q_n$
whose vertex set is $\{0,1\}^n$, the set of all binary strings of length $n$,
and whose edge set is the exactly the set of pairs of strings that differ by exactly
one bit. Small examples are illustrated in Fig.~3. (As an exercise, try to figure out
which strings correspond to which vertices!)
\midinsert
$$\epsfbox{qn.ps}$$
\vskip5pt
\caption{Hypercubes $Q_n$ for small $n$.}
\endinsert
\goodbreak

@@

\medskip\boldlabel Walks, paths, and cycles.
A {\it walk} in $G = (V,E)$ is a sequence $\sigma$ of vertices
$$\sigma = (v_0, v_1, \ldots, v_n)$$
such that $v_i v_{i+1}\in E$ for all $0\le i \le n-1$. The {\it endpoints}
of $\sigma$ are $v_0$ and $v_n$, and the {\it length} of $\sigma$, denoted $|\sigma|$,
is the number of edges traversed, namely, $n$.
The walk is said to be {\it closed} if $v_0 = v_n$ and {\it open} otherwise.
A walk is called a {\it path} if no vertex is repeated.

@@

We conclude that $\sigma$ is a cycle.\slug
\bye

@@

Another characterisation of trees is that they are exactly the connected graphs on $n$ vertices
with $n-1$ edges. To prove this, we first need a lemma.
